Cross-Loop Optimization of Arithmetic Intensity for Finite Element Local Assembly

### 2014-09-09 TACO-paper-v2/coffee.tex
2-column layout for listing 4
minor fix

% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

@@

During the assembly phase, the solution of the PDE is approximated by
executing a problem-specific kernel over all cells, or elements, in the
discretized domain. In this work, we focus on relatively low order finite
element methods, in which an assembly kernel's working set is usually small
enough to fit the L1 cache. Low order methods are by no means exotic: they
are employed in a wide variety of fields, including climate and ocean
modeling, computational fluid dynamics, and structural mechanics. The
efficient assembly of high order methods such as the spectral element
method \cite{SPENCER} requires a significantly different loop nest structure.
High order methods are therefore excluded from our study.

@@

\KwSty{void} helmholtz(\KwSty{double} A[4][4], \KwSty{double} **coords) $\lbrace$\\
~~// K, det = Compute Jacobian (coords) \\
~~// Declaration of basis function matrices \\
~~\\
~~\KwSty{for} (\KwSty{int} i = 0; i$<$3; i++) $\lbrace$ \\
~~~~// Do generalized loop-invariant code motion \\
~~~~\KwSty{for} (\KwSty{int} j = 0; j$<$4; j+=4) \\
~~~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$4; k+=4) $\lbrace$\\
~~~~~~~~// \emph{load} and \emph{set} intrinsics \\
~~~~~~~~// Compute A[0,0], A[1,1], A[2,2], A[3,3] \\
~~~~~~~~// One \emph{permute\_pd} intrinsic per \texttt{k}-loop \emph{load}\\
~~~~~~~~// Compute A[0,1], A[1,0], A[2,3], A[3,2] \\
~~~~~~~~// One \emph{permute2f128\_pd} intrinsic per \texttt{k}-loop \emph{load}\\
~~~~~~~~// ...\\
~~~~~~$\rbrace$\\
~~~~// Remainder loop (from $j=4$ to $j=6$)\\
~~$\rbrace$\\
~~// Restore the storage layout:\\
~~\KwSty{for} (\KwSty{int} j = 0; j$<$4; j+=4) $\lbrace$\\
~~~~\KwSty{$\_\_$m256d} r0, r1, r2, r3, r4, r5, r6, r7;\\
~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$4; k+=4) $\lbrace$\\
~~~~~~r0 = $\_$mm256$\_$load$\_$pd ($\&$A[j+0][k]);\\
~~~~~~// Load A[j+1][k], A[j+2][k], A[j+3][k]\\
~~~~~~r4 = $\_$mm256$\_$unpackhi$\_$pd (r1, r0);\\
~~~~~~r5 = $\_$mm256$\_$unpacklo$\_$pd (r0, r1);\\
~~~~~~r6 = $\_$mm256$\_$unpackhi$\_$pd (r2, r3);\\
~~~~~~r7 = $\_$mm256$\_$unpacklo$\_$pd (r3, r2);\\
~~~~~~r0 = $\_$mm256$\_$permute2f128$\_$pd (r5, r7, 32);\\
~~~~~~r1 = $\_$mm256$\_$permute2f128$\_$pd (r4, r6, 32);\\
~~~~~~r2 = $\_$mm256$\_$permute2f128$\_$pd (r7, r5, 49);\\
~~~~~~r3 = $\_$mm256$\_$permute2f128$\_$pd (r6, r4, 49);\\
~~~~~~$\_$mm256$\_$store$\_$pd ($\&$A[j+0][k], r0);\\
~~~~~~// Store A[j+1][k], A[j+2][k], A[j+3][k]\\
~~~~$\rbrace$\\
~~$\rbrace$\\
$\rbrace$
\caption{Local assembly code generated by Firedrake for the Helmholtz
  problem after application of \emph{op-vect} on top of the
  optimizations shown in Listing~\ref{code:helmholtz-licm}. In this
  example, the unroll-and-jam factor is 1.}
\label{code:helmholtz-opvect}
\end{algorithm}

>>>>

% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}
\usepackage{multicol}

@@

During the assembly phase, the solution of the PDE is approximated by
executing a problem-specific kernel over all cells, or elements, in the
discretized domain. In this work, we focus on relatively low order finite
element methods, in which an assembly kernel's working set is usually small
enough to fit the L1 cache of traditional CPU architectures. Low order methods are by no means exotic: they
are employed in a wide variety of fields, including climate and ocean
modeling, computational fluid dynamics, and structural mechanics. The
efficient assembly of high order methods such as the spectral element
method \cite{SPENCER} requires a significantly different loop nest structure.
High order methods are therefore excluded from our study.

@@

\begin{multicols}{2}
\KwSty{void} helmholtz(\KwSty{double} A[4][4], \KwSty{double} **coords) $\lbrace$\\
~~// K, det = Compute Jacobian (coords) \\
~~// Declaration of basis function matrices \\
~~\\
~~\KwSty{for} (\KwSty{int} i = 0; i$<$3; i++) $\lbrace$ \\
~~~~// Do generalized loop-invariant code motion \\
~~~~\KwSty{for} (\KwSty{int} j = 0; j$<$4; j+=4) \\
~~~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$4; k+=4) $\lbrace$\\
~~~~~~~~// \emph{load} and \emph{set} intrinsics \\
~~~~~~~~// Compute A[0,0], A[1,1], A[2,2], A[3,3] \\
~~~~~~~~// One \emph{permute\_pd} intrinsic per \texttt{k}-loop \emph{load}\\
~~~~~~~~// Compute A[0,1], A[1,0], A[2,3], A[3,2] \\
~~~~~~~~// One \emph{permute2f128\_pd} intrinsic per \texttt{k}-loop \emph{load}\\
~~~~~~~~// ...\\
~~~~~~$\rbrace$\\
~~~~// Remainder loop (from $j=4$ to $j=6$)\\
~~$\rbrace$\\
~~// Restore the storage layout:\\
~~\KwSty{for} (\KwSty{int} j = 0; j$<$4; j+=4) $\lbrace$\\
~~~~\KwSty{$\_\_$m256d} r0, r1, r2, r3, r4, r5, r6, r7;\\
~~~~\KwSty{for} (\KwSty{int} k = 0; k$<$4; k+=4) $\lbrace$\\
~~~~~~r0 = $\_$mm256$\_$load$\_$pd ($\&$A[j+0][k]);\\
~~~~~~// Load A[j+1][k], A[j+2][k], A[j+3][k]\\
~~~~~~r4 = $\_$mm256$\_$unpackhi$\_$pd (r1, r0);\\
~~~~~~r5 = $\_$mm256$\_$unpacklo$\_$pd (r0, r1);\\
~~~~~~r6 = $\_$mm256$\_$unpackhi$\_$pd (r2, r3);\\
~~~~~~r7 = $\_$mm256$\_$unpacklo$\_$pd (r3, r2);\\
~~~~~~r0 = $\_$mm256$\_$permute2f128$\_$pd (r5, r7, 32);\\
~~~~~~r1 = $\_$mm256$\_$permute2f128$\_$pd (r4, r6, 32);\\
~~~~~~r2 = $\_$mm256$\_$permute2f128$\_$pd (r7, r5, 49);\\
~~~~~~r3 = $\_$mm256$\_$permute2f128$\_$pd (r6, r4, 49);\\
~~~~~~$\_$mm256$\_$store$\_$pd ($\&$A[j+0][k], r0);\\
~~~~~~// Store A[j+1][k], A[j+2][k], A[j+3][k]\\
~~~~$\rbrace$\\
~~$\rbrace$\\
$\rbrace$
\end{multicols}
\caption{Local assembly code generated by Firedrake for the Helmholtz
  problem after application of \emph{op-vect} on top of the
  optimizations shown in Listing~\ref{code:helmholtz-licm}. In this
  example, the unroll-and-jam factor is 1.}
\label{code:helmholtz-opvect}
\end{algorithm}
